---
title: "Project_02_RMD"
author: "David Suffolk"
date: "4/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATASET 1 ##  
## CROW SAMPLING ##

### Data and Libraries ###  
  
#### Libraries ####  
  
```{r}
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(usmap)
library(rpart)
library(rpart.plot)
```
  
#### Import data file ####  
  
```{r}
data <- read.csv("./Project 02/nwCrow_bloodParasites_alaska_smith_2007_2008/nwCrow_sampling_alaska_smith_2007_2008.csv")
head(data)
```
  
#### Summary of Data ####  
  
```{r}
summary(data)
```

### Histograms ###  

#### Sex ####  
1 = Male  
2 = Female  
  
```{r}
ggplot(data, aes(x=SEX)) + geom_histogram(binwidth=0.75, color="black", fill="white")
```

#### Location ####  
  
```{r}
ggplot(data, aes(x=LOC)) + geom_histogram(binwidth=0.75, color="black", fill="white", stat="count")
```

#### Location ####  
  
```{r}
ggplot(data, aes(x=AGE)) + geom_histogram(binwidth=0.75, color="black", fill="white")
```

#### AKD ####  

Designation given to sampled Northwestern crows based on presence or absence of clinical signs of Avian Keratin Disorder.  
1 = ABSENT  
0 = PRESENT  
  
```{r}
ggplot(data, aes(x=AKD)) + geom_histogram(binwidth=0.75, color="black", fill="white")
```
  
#### TARSUS ####  

Diagonal tarsus length (mm) of sampled Northwestern crows
  
```{r}
ggplot(data, aes(x=TARSUS)) + geom_histogram(binwidth=0.75, color="black", fill="white")
```

#### WING ####  

Unflattened wing chord length (mm) of sampled Northwestern crows
  
```{r}
ggplot(data, aes(x=WING)) + geom_histogram(bins=30, color="black", fill="white")
```

#### MASS ####  

Mass (g) of sampled Northwestern crows. Blank cells indicate that mass was not measured for that individual
  
```{r}
ggplot(data, aes(x=MASS)) + geom_histogram(bins=30, color="black", fill="white")
```

## DATASET 2 ##  
## BLOOD PARASITES ##  
  
#### Import data file ####  
  
```{r}
data_2 <- read.csv("./Project 02/nwCrow_bloodParasites_alaska_smith_2007_2008/nwCrow_bloodParasites_alaska_smith_2007_2008.csv")
head(data_2)
```
  
#### Summary of Data ####  
  
```{r}
summary(data_2)
```

### Data Cleanup ###

```{r}
#filter(data_2, data_2$LEUC1 == 0 & data_2$LEUC2 == 0)
```

### Histograms ###  
  
#### LEUC1 ####  

Result from the first replicate of the nested-PCR test for Leucocytozoon parasite infection  
0 = Negative result: nested-PCR test resulted in no product, single-stranded sequences, or products that could not be assigned via NCBI nucleotide BLAST  
1 = Positive result: nested-PCR resulted in double-stranded, target mtDNA sequence that was successfully assigned to Leucocytozoon parasite genera via NCBI nucleotide BLAST  
  
```{r}
ggplot(data_2, aes(x=LEUC1)) + geom_histogram(binwidth = 0.75, color="black", fill="white")
```
  
#### LEUC2 ####  

Result from the second replicate of the nested-PCR test for Leucocytozoon parasite infection  
0 = Negative result: nested-PCR test resulted in no product, single-stranded sequences, or products that could not be assigned via NCBI nucleotide BLAST  
1 = Positive result: nested-PCR resulted in double-stranded, target mtDNA sequence that was successfully assigned to Leucocytozoon parasite genera via NCBI nucleotide BLAST  
  
```{r}
ggplot(data_2, aes(x=LEUC2)) + geom_histogram(binwidth = 0.75, color="black", fill="white")
```
  
#### HAEM1 ####  
  
Result from the first replicate of the nested-PCR test for Haemoproteus parasite infection  
0 = Negative result: nested-PCR test resulted in no product, single-stranded sequences, or products that could not be assigned via NCBI nucleotide BLAST  
1 = Positive result: nested-PCR resulted in double-stranded, target mtDNA sequence that was successfully assigned to Haemoproteus parasite genera via NCBI nucleotide BLAST  
  
```{r}
ggplot(data_2, aes(x=HAEM1)) + geom_histogram(binwidth = 0.75, color="black", fill="white")
```
  
#### HAEM2 ####  
  
Result from the second replicate of the nested-PCR test for Haemoproteus parasite infection  
0 = Negative result: nested-PCR test resulted in no product, single-stranded sequences, or products that could not be assigned via NCBI nucleotide BLAST   
1 = Positive result: nested-PCR resulted in double-stranded, target mtDNA sequence that was successfully assigned to Haemoproteus parasite genera via NCBI nucleotide BLAST  
  
```{r}
ggplot(data_2, aes(x=HAEM2)) + geom_histogram(binwidth = 0.75, color="black", fill="white")
```
  
#### PLAS1 ####  
  
Result from the first replicate of the nested-PCR test for Plasmodium parasite infection  
0 = Negative result: nested-PCR test resulted in no product, single-stranded sequences, or products that could not be assigned via NCBI nucleotide BLAST  
1 = Positive result: nested-PCR resulted in double-stranded, target mtDNA sequence that was successfully assigned to Plasmodium parasite genera via NCBI nucleotide BLAST  
  
```{r}
ggplot(data_2, aes(x=PLAS1)) + geom_histogram(binwidth = 0.75, color="black", fill="white")
```
  
#### PLAS2 ####  
  
Result from the second replicate of the nested-PCR test for Plasmodium parasite infection  
0 = Negative result: nested-PCR test resulted in no product, single-stranded sequences, or products that could not be assigned via NCBI nucleotide BLAST  
1 = Positive result: nested-PCR resulted in double-stranded, target mtDNA sequence that was successfully assigned to Plasmodium parasite genera via NCBI nucleotide BLAST 
  
```{r}
ggplot(data_2, aes(x=PLAS2)) + geom_histogram(binwidth = 0.75, color="black", fill="white")
```
  
#### Leuc_GenBank_Accession ####  
  
GenBank accession number for Leucocytozoon cytochrome b (cyt b) mitochondrial DNA (mtDNA) target fragment following a positive result from any of two replicates of the nested-PCR. Blank cells signify negative results from all replicates of the nested-PCR  
  
```{r}
ggplot(data_2, aes(x=Leuc_GenBank_Accession)) + geom_histogram(binwidth = 0.75, color="black", fill="white", stat="count")
```
  
#### Haem_GenBank_Accession ####  
  
GenBank accession number for Haemoproteus cytochrome b (cyt b) mitochondrial DNA (mtDNA) target fragment following a positive result from any of two replicates of the nested-PCR. Blank cells signify negative results from all replicates of the nested-PCR  
  
```{r}
ggplot(data_2, aes(x=Haem_GenBank_Accession)) + geom_histogram(binwidth = 0.75, color="black", fill="white", stat="count")
```
  
#### Plas_GenBank_Accession ####  
  
GenBank accession number for Plasmodium cytochrome b (cyt b) mitochondrial DNA (mtDNA) target fragment following a positive result from any of two replicates of the nested-PCR. Blank cells signify negative results from all replicates of the nested-PCR  
  
```{r}
ggplot(data_2, aes(x=Plas_GenBank_Accession)) + geom_histogram(binwidth = 0.75, color="black", fill="white", stat="count")
```
  
### Merge Datasets ###  
  
```{r}
de <- merge(data, data_2, by=0, all=TRUE)
head(de)
```
  
### Covariance ###  
  
Create variables
  
```{r}
SEX = de$SEX
LOC = de$LOC
AGE = de$AGE
AKD = de$AKD
TARSUS = de$TARSUS
WING = de$WING
MASS = de$MASS
LEUC1 = de$LEUC1
HAEM1 = de$HAEM1
PLAS1 = de$PLAS1
```  
  
#### AKD ####  
  
```{r}
cov(SEX,AKD)
ggplot(de, aes(x=SEX, y=AKD)) +
  geom_point(size=2, shape=23)
```
  
```{r}
cov(AGE,AKD)
```
  
```{r}
cov(TARSUS,AKD)
```
  
```{r}
cov(WING,AKD)
```

```{r}
cov(MASS,AKD)
```
  
#### LEUC1 ####
  
```{r}
cov(SEX,LEUC1)
```
  
```{r}
cov(AGE,LEUC1)
```
  
```{r}
cov(TARSUS,LEUC1)
```
  
```{r}
cov(WING,LEUC1)
```

```{r}
cov(MASS,LEUC1)
```
  
#### HAEM1 ####  
  
```{r}
cov(SEX,HAEM1)
```
  
```{r}
cov(AGE,HAEM1)
```
  
```{r}
cov(TARSUS,HAEM1)
```
  
```{r}
cov(WING,HAEM1)
```

```{r}
cov(MASS,HAEM1)
```


#### PLAS1 ####  

```{r}
cov(SEX,PLAS1)
```
  
```{r}
cov(AGE,PLAS1)
```
  
```{r}
cov(TARSUS,PLAS1)
```
  
```{r}
cov(WING,PLAS1)
```

```{r}
cov(MASS,PLAS1)
```  
  
  
### Correlation Coefficient ###   
  
#### AKD ####  
  
```{r}
cor(SEX,AKD)
```
  
```{r}
cor(AGE,AKD)
```
  
```{r}
cor(TARSUS,AKD)
```
  
```{r}
cor(WING,AKD)
```

```{r}
cor(MASS,AKD)
```
  
#### LEUC1 ####
  
```{r}
cor(SEX,LEUC1)
```
  
```{r}
cor(AGE,LEUC1)
```
  
```{r}
cor(TARSUS,LEUC1)
```
  
```{r}
cor(WING,LEUC1)
```

```{r}
cor(MASS,LEUC1)
```
  
#### HAEM1 ####  
  
```{r}
cor(SEX,HAEM1)
```
  
```{r}
cor(AGE,HAEM1)
```
  
```{r}
cor(TARSUS,HAEM1)
```
  
```{r}
cor(WING,HAEM1)
```

```{r}
cor(MASS,HAEM1)
```


#### PLAS1 ####  

```{r}
cor(SEX,PLAS1)
```
  
```{r}
cor(AGE,PLAS1)
```
  
```{r}
cor(TARSUS,PLAS1)
```
  
```{r}
cor(WING,PLAS1)
```

```{r}
cor(MASS,PLAS1)
```    
  
  
# AKD and Pathogens - Correlation Coefficients ###
```{r}
cor(AKD,LEUC1)
``` 

```{r}
cor(AKD,HAEM1)
``` 

```{r}
cor(AKD,PLAS1)
```   
  
# AKD and Pathogens - Covariance ###
```{r}
cov(AKD,LEUC1)
``` 

```{r}
cov(AKD,HAEM1)
``` 

```{r}
cov(AKD,PLAS1)
```  
  
  
### LM on LEUC1 ###  
  
```{r}
LEUC1_LM <- lm(LEUC1~LOC+SEX+AGE+AKD+TARSUS+WING+MASS+HAEM1+PLAS1, data=de)
summary(LEUC1_LM)
```
  
Positive P values:  
Most are negative or close to zero  
  
### LM on HAEM1 ###  
  
```{r}
HAEM1_LM <- lm(HAEM1~LOC+SEX+AGE+AKD+TARSUS+WING+MASS+LEUC1+PLAS1, data=de)
summary(HAEM1_LM)
```
  
Positive P values:  
Most are negative or close to zero  
  
### LM on PLAS1 ###  
  
```{r}
PLAS1_LM <- lm(PLAS1~LOC+SEX+AGE+AKD+TARSUS+WING+MASS+LEUC1+HAEM1, data=de)
summary(PLAS1_LM)
```
  
Positive P values:  
Most are negative or close to zero 
  
### LM on AKD ###  
  
```{r}
AKD_LM <- lm(AKD~LOC+SEX+AGE+PLAS1+TARSUS+WING+MASS+LEUC1+HAEM1, data=de)
summary(AKD_LM)
```
  
Positive P values:  
Most are negative or close to zero 
  
### LM on AKD with Pathogens ###  
  
```{r}
AKD_LM_2 <- lm(AKD~PLAS1+LEUC1+HAEM1, data=de)
summary(AKD_LM_2)
```
  
### LM on PLAS1 with AKD and Pathogens ###  
  
```{r}
PLAS1_LM_2 <- lm(PLAS1~AKD+LEUC1+HAEM1, data=de)
summary(PLAS1_LM_2)
```
 
### LM on LEUC1 with AKD and Pathogens ###  
  
```{r}
LEUC1_LM_2 <- lm(LEUC1~PLAS1+AKD+HAEM1, data=de)
summary(LEUC1_LM_2)
```

### LM on HAEM1 with AKD and Pathogens ###  
  
```{r}
HAEM1_LM_2 <- lm(HAEM1~PLAS1+LEUC1+AKD, data=de)
summary(HAEM1_LM_2)
```
  
